---
title: "Statistical Analysis"
author: "Agatha North"
date: "December 3, 2017"
output: html_document
---

```{r include=FALSE}
require(tidyverse)
load("myData.RData")
```


# Analysis and Results

When digging into the data certain things quickly became apparent. First and foremost was that the size of the companies involved in the data varied, and as a result, the volume of their votes, likes, dislikes, comments, and other data of relevance. 

```{r echo=TRUE}
ggplot(votes, aes(vote, fill = companyAlias)) + geom_bar() + facet_grid(~ votes$companyAlias) + labs(title = "Votes by Company")

```

Some things to note from the plot above, companies 2 and 12 are particularly happy, with the strong majority of their votes marked as 4. Companies 1, 23, and 31 also stand out in terms of overall happiness compared to unhappiness. On the otherhand company 23 stands out as the most evenly distributed between negative and positve votes, which incidentally also makes it the least happy of the batch.

```{r echo=TRUE}
ggplot(comments, aes(likes, fill = companyAlias)) + geom_histogram(binwidth = 1) + labs(title ="Likes by Company")
```
```{r echo=TRUE}
ggplot(comments, aes(dislikes, fill = companyAlias)) + geom_histogram(binwidth = 1) + labs(title = "Dislikes by Company")
```
It was far more common for comments to receive likes than dislikes in every company. These slopes were overall unsurpising.

```{r echo=TRUE}
ggplot(EmpCombined, aes(EmpID, col = stillExists, x = MeanVote, y = VoteCount)) + geom_point(position = "jitter", alpha=0.6)
```
Investigating the individual votes when compared to how often an employee voted was quite fruitful. The distribution of votes for those who churned was extremely even across the mean vote, but the density of all votes trended towards the higher end. However, I did not expect to see individuals who's mean vote was roughly 1 have anywhere near the same vote count as individuals with a higher level of day to day happiness. Clearly day to day happiness wasn't all that strong of an indicator on its own, which proved useful later. It's facinating to think that there were some individuals who publically voted their unhappiness for over half a year and still remained at their company. 

```{r echo=TRUE}
ggplot(EmpCombined, aes(EmpID, col = stillExists, x=VoteCount, y=AvgLength)) + geom_point(position = "jitter", alpha=0.4)
```

```{r echo=TRUE}
ggplot(EmpCombined, aes(EmpID, col = stillExists, x=VoteCount, y=AvgDislikes)) + geom_point(position = "jitter", alpha=0.4)
```

```{r echo=TRUE}
ggplot(EmpCombined, aes(EmpID, col = stillExists, x=VoteCount, y=AvgLikes)) + geom_point(position = "jitter", alpha=0.4)
```

Looking at the average length, average dislikes, and average likes of individuals compared to their vote count, it's immediately clear that these three elements of engagement were strong indicators of investment in the app, if not the company as well. Interestingly there were few employees that seemed to be engaged by both the votes and also the comments.
